---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)


```

# absmapsdata

The `absmapsdata` package exists to make it easier to produce maps from ABS data in R. The package contains compressed, tidied, and lazily-loadable `sf` objects containing geometric information about ABS data structures.

Before we get into the 'what problem is this package solving' details, let's look at some examples so that you can copy-paste into your own console and replicate out-of-the-box. 


## Installation

You can install `absmapsdata` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("wfmackey/absmapsdata")
```

We will use the `sf` and `tidyverse` packages in this example:

```{r, message = FALSE}
library(tidyverse)
library(sf)
```


## Just show me how to make a map with this package

### Using the package's pre-loaded data

The `absmapsdata` package comes with pre-downloaded and pre-processed data. To load a particular geospatial object, simply load the package and call the object:

```{r}

library(absmapsdata)

mapdata1 <- sa32011

glimpse(mapdata1)
```

Or

```{r}
mapdata2 <- sa22016

glimpse(mapdata2)
```


The resulting `sf` object contains one observation per area (in the following examples, one observation per `sa3`). It stores the geometry information in the `geometry` variable, which is a nested list describing the area's polygon. The object can be joined to a standard `data.frame` or `tibble` and can be used with `dplyr` functions.


### Creating maps with your `sf` object
We do all this so we can create gorgeous maps. And with the `sf` object in hand, plotting a map via `ggplot` and `geom_sf` is simple.

```{r}
map <-
sa32016 %>%
  filter(gcc_name_2016 == "Greater Melbourne") %>%   # let's just look Melbourne
  ggplot() +
  geom_sf(aes(geometry = geometry))  # use the geometry variable

map
```

The data also include centorids of each area, and we can add these points to the map with the `cent_lat` and `cent_long` variables using `geom_point`.

```{r}
map <-
sa32016 %>%
  filter(gcc_name_2016 == "Greater Melbourne") %>%   # let's just look Melbourne
  ggplot() +
  geom_sf(aes(geometry = geometry)) +   # use the geometry variable
  geom_point(aes(cent_lat, cent_long))  # use the centroid lat and longs

map
```

Cool. But, sidenote, this all looks a bit ugly. We can pretty it up using `ggplot` tweaks. See the comments on each line for its objective. Also note that we're filling the areas by their `areasqkm` size, another variable included in the `sf` object (we'll replace this with more interesting data in the next section).

```{r}
map <-
sa32016 %>%
  filter(gcc_name_2016 == "Greater Melbourne") %>%   # let's just look Melbourne
  ggplot() +
  geom_sf(aes(geometry = geometry,  # use the geometry variable
              fill = areasqkm_2016),     # fill by area size
          lwd = 0,                  # remove borders
          show.legend = FALSE) +    # remove legend
  geom_point(aes(cent_lat,
                 cent_long),        # use the centroid lat and longs
             colour = "white") +    # make the points white
  theme_void() +                    # clears other plot elements
  coord_sf(datum = NA)              # fixes a gridline bug in theme_void()

map
```

```{r, test label, eval = FALSE, include = FALSE}

data <- sa42016 %>% 
  mutate(label = gsub("Melbourne\\s\\-", "", sa4_name_2016),
         label = gsub("\\s", "\n", label))

data %>%
  filter(gcc_name_2016 == "Greater Melbourne") %>%   # let's just look Melbourne
  ggplot() +
  geom_sf(aes(geometry = geometry,  # use the geometry variable
              fill = areasqkm_2016),     # fill by area size
          lwd = 0,                  # remove borders
          show.legend = FALSE) +    # remove legend
  geom_text(aes(cent_lat, cent_long, label = label),
            size = 3,
            colour = "white") +
  theme_void() +                    # clears other plot elements
  coord_sf(datum = NA)              # fixes a gridline bug in theme_void()

```


## Joining with other datasets

At some point, we'll want to join our spatial data with data-of-interest. The variables in our mapping data---stating the numeric code and name of each area and parent area---will make this _relatively_ easy.

For example: suppose we had a simple dataset of median income by SA3 over time. 

```{r, include = FALSE}
income <- read_csv("vignettes/data/median_income_sa3.csv")
```


```{r, eval = FALSE}
# Read data in
income <- read_csv("data/median_income_sa3.csv")
```

```{r}
# This data contains a variable of sa4 which 
# can be joined to the sa4_name_2016 variable:
combined_data <- income %>% 
  left_join(sa32016, by = "sa3_name_2016")
```

Now that we have a tidy dataset with 1) the income data we want to plot, and 2) the geometry of the areas, we can plot income by area:

```{r}
map <-
combined_data %>%
  filter(gcc_name_2016 == "Greater Melbourne") %>%   # let's just look Melbourne
  ggplot() +
  geom_sf(aes(geometry = geometry,  # use the geometry variable
              fill = median_income),        # fill by unemployment rate
          lwd = 0) +                # remove borders
  theme_void() +                    # clears other plot elements
  coord_sf(datum = NA) +            # fixes a gridline bug in theme_void()
  labs(fill = "Median income")

map
```




## Why does this package exist?

The motivation for this package is that maps are cool and fun and are, sometimes, the best way to communicate data.
And making maps is `R` with `ggplot` is relatively easy _when you have the right `object`_.

Getting the right `object` is not technically difficult, but requires research into the best-thing-to-do at each of the following steps:

*  Find the ASGS ABS spatial-data page and determine the right file to download.
*  Read the shapefile into `R` using one-of-many import tools.
*  Convert the object into something usable.
*  Clean up any inconsistencies and apply consistent variable naming/values across areas and years.
*  Find an appropriate compression function and level to optimise output.

For me, at least, finding the correct information and developing the best set of steps was a little bit interesting but mostly tedious and annoying. The `absmapsdata` package holds this data for you, so you can spend more time making maps, and less time on Stack Overflow, the ABS website, and [lovely-people's wonderful blogs](https://www.neonscience.org/dc-open-shapefiles-r).


## How does this package do the-things-it-does

The `absmapsdata` package simple holds compressed and easy to use data (`sf` objects) for you use.

It is a data-only-based sibling of `absmaps`, which holds functionality to download more shapefile data from the ABS and compress it to a level you desire. However, this comes at a cost: the mapping software that `absmaps` is built on can be a bit fiddly to install.

If you would like to do these things, please feel free to install `absmapsdata`. 

## I want to complain about this package

Fair enough! The best avenue is via a Github issue at (wfmackey/absmapsdata). This is also the best place to request data that isn't yet available in the package.
